---
title: "Getting Started with SuperIntervals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with SuperIntervals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

SuperIntervals provides fast interval overlap queries through the IntervalMap data structure. This is particularly useful for genomic analysis, time series data, and any application requiring efficient range-based lookups.

## Basic Usage

```{r setup}
library(superintervalsr)
```

### Creating and Building an IntervalMap

```{r basic}
# Create a new IntervalMap
imap <- IntervalMap()

# Add some intervals with associated values
add(imap, 10, 20, "Gene A")
add(imap, 15, 25, "Gene B")
add(imap, 30, 40, "Gene C")

# Build the index (required before queries)
build(imap)

# View the IntervalMap
print(imap)
```

### Searching for Overlaps

```{r search}
# Find all values that overlap with range 12-18
values <- search_values(imap, 12, 18)
print(values)

# Get the intervals (start, end) that overlap
keys <- search_keys(imap, 12, 18)
print(keys)

# Get complete information (start, end, value)
items <- search_items(imap, 12, 18)
print(items)
```

### Query Statistics

```{r stats}
# Check if any intervals overlap a range
has_overlaps(imap, 5, 12)
has_overlaps(imap, 50, 60)

# Count overlapping intervals
count(imap, 12, 18)

# Get coverage statistics
coverage(imap, 10, 40)
```

### Accessing Individual Intervals

```{r access}
# Get interval at specific index
at(imap, 1)

# Get just the start position
starts_at(imap, 1)

# Get just the end position
ends_at(imap, 1)

# Get just the value
data_at(imap, 1)
```

## Performance Tips

- Call `build()` after adding all intervals and before performing queries
- Use `reserve()` if you know the number of intervals in advance
- The IntervalMap is optimized for many queries on the same set of intervals

## Use Cases

SuperIntervals is ideal for:

- **Genomics**: Finding genes/features overlapping genomic regions
- **Time series**: Querying events within time ranges
- **Spatial analysis**: Range-based spatial queries
- **Data analysis**: Any scenario requiring fast interval overlap detection